---
title: "Study of Cardiological Patients: Preconditions and Procedures"
author: "Group 8"
date: "`r Sys.Date()`"
output: rmdformats::material
fontsize: 9pt
---
# Introduction

This report aims to, through medical data analysis, give an understanding of the dataset that was given to the team, and to use and train a ML model in order to predict what complications patients may present after different medical procedures as accurately as possible.

The dataset used in this assignment was given by Austral University Hospital's cardiology department, located in Pilar, Buenos Aires.

This document is organized into three sections: Data Analysis, which shows generalized queries on the dataset; Interpretation, which seeks to give further understanding of the given dataset in its context, and Predictor, where the engineering behind the model is explained.

To analyze the dataset in the correct way and extract the richest information we started with researching the variables meaning, medical terms included, and cardiological procedures from different information sources. We then proceeded with the implementation of data analysis tools and strategies to fulfill our objectives.

#  Data Analysis 
<font size="2">In this section we evaluate different variables in the dataset and show graphs.
</font>
<p>&nbsp;</p>
<font size="5"> **Sex and Age** </font>
```{r, include = FALSE, results='asis'}
library(plotrix)
library(plotly)
library(ggplot2)
library(ggridges)
library(dplyr)
library(hrbrthemes)
library(Hmisc)
library(XML)
library(reshape2)
library(plyr)
library(ggplot2)
library(viridis)
library(rmdformats)

dataset <- read.csv("dataset.csv", sep=";")
x <- cut2(dataset$EDAD, seq(10,90,5))

yes <- dataset$SEXO
displayIngresoSegunEdad<- table(dataset$SEXO,dataset$EDAD)
displayIngresoSegunEdad
textIntervals <- as.character(x)
f <- as.data.frame(table(x))

df <- data.frame(textIntervals, yes)

counts <- ddply(df, .(df$textIntervals, df$yes), nrow)
names(counts) <- c("ageInterval", "Sexo", "Freq")
counts

## pyramid charts are two barcharts with axes flipped

counts$Freq <- ifelse(counts$Sexo == "MASC", -1*counts$Freq, counts$Freq)

pyramid <- ggplot(counts, aes(x = counts$ageInterval, y = counts$Freq, fill = counts$Sexo)) + 
  geom_bar(data = subset(counts, counts$Sexo == "Female"), stat = "identity") +
  geom_bar(data = subset(counts, counts$Sexo == "Male"), stat = "identity") +
  scale_y_continuous(breaks = seq(-60, 60, 10),labels = paste0(as.character(c(seq(60, 0, -10), seq(10, 60, 10))),'')) + 
  coord_flip() + 
  xlab("Age") + ylab("Number Of Patients")+
  scale_fill_manual(name = "Sex", labels = c("Female", "Male"),values = c("#f08080","#4ea9ca"))+
  theme_bw()

```

```{r fig.width=6, fig.height=6, fig.align='center',echo=FALSE}
pyramid
```

<font size="2">This pyramid graph allows to observe the sex distribution by age. There is a much higher percentage of men in the dataset, which are mainly concentrated between 60 and 70 years old.</font>
<p>&nbsp;</p>
<font size="5.5"> **Reason for Admission** </font>

```{r, include = FALSE, results='asis'}
displayIngresoSegunEdad<- table(dataset$MOTIVO.DE.INGRESO,dataset$EDAD)
motivoIngreso<- factor(dataset$MOTIVO.DE.INGRESO)
motivoIngreso

yes <- dataset[which(dataset$MOTIVO.DE.INGRESO == "PROG CORONARIO" | dataset$MOTIVO.DE.INGRESO == "PROG VALV" | dataset$MOTIVO.DE.INGRESO == "SCA no SST"),]
yes
x <- cut2(yes$EDAD, seq(10,90,15))


displayIngresoSegunEdad<- table(yes$MOTIVO.DE.INGRESO,yes$EDAD)
displayIngresoSegunEdad
textIntervals <- as.character(x)
f <- as.data.frame(table(x))

df <- data.frame(textIntervals, yes)

counts <- ddply(df, .(df$textIntervals, df$MOTIVO.DE.INGRESO), nrow)
names(counts) <- c("ageInterval", "Motivo", "Freq")
counts

plot <-ggplot(counts, aes(fill=counts$ageInterval, y=counts$Freq, x=counts$Motivo)) + 
  geom_bar(position="stack", stat="identity")+
  scale_fill_manual(labels = c('[10;25)','[25;40)','[40;55)','[55;70)','[70;85)','[85;92)'),values = c("#4169E1","#015d67","#36a792","#63bd99","#87CEEB","#B0E0E6"))+
  xlab("Reason for Admission") + ylab("Frequency")+
  labs(fill = "Age Intervals")

plot

```
```{r fig.width=8, fig.height=6, fig.align='center',echo=FALSE,warning=FALSE}
plot
```

<font size="2">The aim of this graph is to show the main causes of admission in the Hospital, according to the age of the patients. It is possible to see that Coronary Programmed Operations stands out as the most common reason.</font>
<p>&nbsp;</p>
<font size="5.5"> **Procedures** </font>

```{r, include = FALSE, results='asis'}
datasetSinNA = dataset[which(dataset$PROCEDIMIENTO != "CIRUGIA, ENDOVALVULA"),]
distributionPlot<-datasetSinNA %>%
  ggplot(aes(x= EDAD, fill = PROCEDIMIENTO,..count..*10)) +
  geom_density(alpha = 0.5) +
  theme_ridges() +
  scale_fill_manual(name = "Procedure:", labels = c("Surgery","Angioplasty","Mitral Valve Rep."), values = c("#10217d","#16c2d5","#3CB371"),) +
  theme(legend.position = "top")+
  labs(x="Age", y = "Density")
```
```{r fig.width=6, fig.height=6, fig.align='center',echo=FALSE,warning=FALSE}
distributionPlot
```

<font size="2">This graph shows how procedures are distributed according to the age of the patient. Mitral Valve Replacement far outweighs as the most recurrent procedure in the oldest ages.</font>
<p>&nbsp;</p>
<font size="5.5"> **Types of Surgery** </font>

```{r, include = FALSE, results='asis'}
subdatasetCirugia <- dataset[which(dataset$PROCEDIMIENTO == "CIRUGIA" & (dataset$TIPO.DE.CIRUGIA != "#N/A" & dataset$TIPO.DE.CIRUGIA != "TUMORES DE AURICULA" & dataset$TIPO.DE.CIRUGIA != "TROMBECTOMIA PULMONAR" & dataset$TIPO.DE.CIRUGIA != "PLASTICA TRICUS, OTRA" & dataset$TIPO.DE.CIRUGIA != "PERICARDIECTOMIA" & dataset$TIPO.DE.CIRUGIA != "DOBLE REEMPLAZO VALVULAR")),]
tiposDeCirugias<- factor(subdatasetCirugia$TIPO.DE.CIRUGIA)
tablaTiposDeCirugias <- table(tiposDeCirugias)
tablaTiposDeCirugias

subdatasetCirugia$TIPO.DE.CIRUGIA[subdatasetCirugia$TIPO.DE.CIRUGIA == "RVM"] <- "Myocardical Revascularization"
subdatasetCirugia$TIPO.DE.CIRUGIA[subdatasetCirugia$TIPO.DE.CIRUGIA == "CRM PURA"] <- "PURE CRM"
subdatasetCirugia$TIPO.DE.CIRUGIA[subdatasetCirugia$TIPO.DE.CIRUGIA == "REEMPLAZO VALVULAR PULMONAR"] <- "Pulmonary Valve Replacement"
subdatasetCirugia$TIPO.DE.CIRUGIA[subdatasetCirugia$TIPO.DE.CIRUGIA == "OTRA"] <- "Other"
subdatasetCirugia$TIPO.DE.CIRUGIA[subdatasetCirugia$TIPO.DE.CIRUGIA == "ANEURISMA AORTICO"] <- "Aortic Aneurysm"
subdatasetCirugia$TIPO.DE.CIRUGIA[subdatasetCirugia$TIPO.DE.CIRUGIA == "CIA"] <- "Interatrial Communication"
subdatasetCirugia$TIPO.DE.CIRUGIA[subdatasetCirugia$TIPO.DE.CIRUGIA == "PLASTICA MITRAL"] <- "Mitral Valve Repair"

tiposDeCirugias

subdatasetCRM = dataset[which(dataset$CRM == "true"),]
numCRM = as.numeric(nrow(subdatasetCRM))
numCRMPura = as.numeric(length(factor(which(subdatasetCRM$TIPO.DE.CIRUGIA == "CRM PURA"))))
percentage<- paste(round((numCRMPura/numCRM)*100), "%", sep="") #Porcentaje de CRMS que son CRM puras
percentage
numCirugia = as.numeric(nrow(subdatasetCirugia))
percentage2<- paste(round((numCRM/numCirugia)*100), "%", sep="") #Porcentaje de Cirugias que son CRM
percentage2
```
```{r fig.width=6, fig.height=6, fig.align='center',echo=FALSE,warning=FALSE}
subdatasetCirugia %>% 
  ggplot(aes(y=TIPO.DE.CIRUGIA), position=position_dodge()) + 
  geom_bar( width = 0.85, color = "darkblue", fill = "lightblue") +
  scale_fill_brewer() +
  theme(legend.position="top")+
  ylab("Type of Surgery") +
  xlab("Frequency")
```

<font size="2">In this frequency graph it is clearly seen how CRM and RVAO are the most common types of surgery.</font>
<p>&nbsp;</p> 
<font size="5.5"> **Complications** </font>

```{r, include = FALSE, result = 'asis'}
pacientesATC <- dataset[which(dataset$PROCEDIMIENTO == "ANGIOPLASTIA"),]
tablaAtc = table(pacientesATC$COMPLICACIONES.ANGIOPLASTIA)

prop.table(tablaAtc)

porcentageNoCompAng <- paste(round((tablaAtc[1]/217)*100), "%", sep="")
porcentageCompAng <- paste(round((tablaAtc[2]/217)*100), "%", sep="")

pacientesCRM <- dataset[which(dataset$PROCEDIMIENTO == "CIRUGIA"),]

totalDeCirugia = nrow(pacientesCRM)

cantCompCRM <- nrow(pacientesCRM[which(pacientesCRM$COMPLICACIONES.INMEDIATAS == 1 | pacientesCRM$COMPLICACIONES.TARDIAS == 1),])

cantCompCRM_I <- nrow(pacientesCRM[which(pacientesCRM$COMPLICACIONES.INMEDIATAS == 1 & pacientesCRM$COMPLICACIONES.TARDIAS == 0 ),])
cantCompCRM_T <- nrow(pacientesCRM[which(pacientesCRM$COMPLICACIONES.INMEDIATAS == 0 & pacientesCRM$COMPLICACIONES.TARDIAS == 1),])
cantCompCRM_IyT<- nrow(pacientesCRM[which(pacientesCRM$COMPLICACIONES.INMEDIATAS == 1 & pacientesCRM$COMPLICACIONES.TARDIAS == 1),])

cantidadTotal = cantCompCRM_I + cantCompCRM_IyT + cantCompCRM_T
porComplicaciones = cantidadTotal/totalDeCirugia;
sinComplicaciones = totalDeCirugia - cantidadTotal

tablaCRM = cbind(c(cantidadTotal),c(sinComplicaciones))
tablaCRM
prop.table(tablaCRM)
A <- matrix(nrow=1,ncol=2, c(cantidadTotal,sinComplicaciones), byrow=TRUE)

porcNoCompCrm <- paste(round((as.numeric(A[1,2]/161)*100)), "%", sep="")
porcCompCrm <- paste(round((as.numeric(A[1,1]/161)*100)), "%", sep="")
```
### *Angioplasty*
```{r fig.width=6, fig.height=6, fig.align='center',echo=FALSE,warning=FALSE}
colores <- c("#6495ED", "#CD5C5C")
RegionLabels<-c(porcentageNoCompAng,porcentageCompAng)
pieATC <-pie(tablaAtc, RegionLabels, col = colores,)
legendLabels = c("No Complications","Complications")
legend("topright", legendLabels, cex = 0.8, fill = colores)
```
<font size="2">This pie graph shows how angioplasty is not a procedure that results in numerous complications.</font>
### *Surgery*
```{r fig.width=6, fig.height=6, fig.align='center',echo=FALSE,warning=FALSE}
colores2 <- c("#CD5C5C", "#6495ED")
RegionLabels<-c(porcCompCrm,porcNoCompCrm)
pieSurgery<-pie(tablaCRM, RegionLabels, col = colores2)
legendLabels = c("Complications","No Complications")
legend("topright", legendLabels, cex = 0.8, fill = colores2)
```
<font size="2">In contrast to angioplasty, a significant percentage of surgery patients suffer complications.</font>

# Interpretation 
<font size="2">In this section we look for information from relations and comparisons between variables.</font>
<p>&nbsp;</p>

<font size="5.5"> **Number of injuries according to ACC (Complex Coronary Anatomy)** </font>
```{r, include = FALSE, result = 'asis'}
subsetNoACC <- dataset[which(dataset$ANATOMIA.CORONARIA.COMPLEJA == 0 & 
                               dataset$NUMERO.DE.LESIONES != "#N/A"),]
frecNumLesionesNoACC <- c(as.numeric(length(factor(
  which(subsetNoACC$NUMERO.DE.LESIONES == 0)))), 
  as.numeric(length(factor(which(subsetNoACC$NUMERO.DE.LESIONES == 1)))), 
  as.numeric(length(factor(which(subsetNoACC$NUMERO.DE.LESIONES == 2)))), 
  as.numeric(length(factor(which(subsetNoACC$NUMERO.DE.LESIONES == 3)))),  
  as.numeric(length(factor(which(subsetNoACC$NUMERO.DE.LESIONES == 4)))), 
  as.numeric(length(factor(which(subsetNoACC$NUMERO.DE.LESIONES == 5))))) 

subsetACC <- dataset[which(dataset$ANATOMIA.CORONARIA.COMPLEJA == 1 & 
                             dataset$NUMERO.DE.LESIONES != "#N/A"),]

frecNumLesionesACC <- c(as.numeric(length(factor(which(subsetACC$NUMERO.DE.LESIONES == 0)))), 
                        as.numeric(length(factor(which(subsetACC$NUMERO.DE.LESIONES == 1)))), 
                        as.numeric(length(factor(which(subsetACC$NUMERO.DE.LESIONES == 2)))), 
                        as.numeric(length(factor(which(subsetACC$NUMERO.DE.LESIONES == 3)))),  
                        as.numeric(length(factor(which(subsetACC$NUMERO.DE.LESIONES == 4)))), 
                        as.numeric(length(factor(which(subsetACC$NUMERO.DE.LESIONES == 5))))) 


cols1<- c("0 Injuries", "1 Injury", "2 Injuries", "3 Injuries", "4 Injuries", "5 Injuries")
rows1<- c("ACC", "No ACC")

procedACC_data <- data.frame(frecNumLesionesACC,frecNumLesionesNoACC, row.names = cols1)

colnames(procedACC_data)[1]<-"ACC"
colnames(procedACC_data)[2]<-"No ACC"



```
```{r fig.width=9, fig.height=5, fig.align='center',echo=FALSE,warning=FALSE}
barplot(t(as.matrix(procedACC_data)),beside=TRUE,xlab="Injuries", ylab="Frequency", col=c("#cd5c5c", "#6495ed"),
        legend.text = colnames(procedACC_data),args.legend=list(cex=0.75,x = "topright"),alpha = 0.6)
  
```
<font size="2">In this graph it is possible to appreciate how most of the ACC patients are located in the zone between 3 and 5 injuries, while the rest of the patients are mainly located in the zone between 0 and 2 injuries.</font>
<p>&nbsp;</p>
<font size="5.5"> **Procedure according to affected vessels** </font>

```{r, include = FALSE, result = 'asis'}
subdatasetVasosAfectados = dataset[which(dataset$Resumen.Coronariopatia !="#N/A" & dataset$Resumen.Coronariopatia !="TCI solo"),]
displayVasosProced <- table(subdatasetVasosAfectados$PROCEDIMIENTO,subdatasetVasosAfectados$Resumen.Coronariopatia)
displayVasosProced

subdatasetVasosAfectados
vasosCirugiaAngioplastia <- subdatasetVasosAfectados[which(subdatasetVasosAfectados$PROCEDIMIENTO == "CIRUGIA" | subdatasetVasosAfectados$PROCEDIMIENTO == "ANGIOPLASTIA"),]
unVaso<-vasosCirugiaAngioplastia[which(vasosCirugiaAngioplastia$Resumen.Coronariopatia == "1 vaso"),]
dosVasos<-vasosCirugiaAngioplastia[which(vasosCirugiaAngioplastia$Resumen.Coronariopatia == "2 vasos"),]
tresVasos<-vasosCirugiaAngioplastia[which(vasosCirugiaAngioplastia$Resumen.Coronariopatia == "3 vasos"),]

coronopatia<- c(c("1 vessel", "1 vessel"), c("2 vessels", "2 vessels"),c("3 vessels", "3 vessels"))
procedimiento<- c("Surgery", "Angioplasty","Surgery", "Angioplasty","Surgery", "Angioplasty")
values<- c(nrow(unVaso[which(unVaso$PROCEDIMIENTO == "CIRUGIA"),]),nrow(unVaso[which(unVaso$PROCEDIMIENTO == "ANGIOPLASTIA"),]),nrow(dosVasos[which(dosVasos$PROCEDIMIENTO == "CIRUGIA"),]),nrow(dosVasos[which(dosVasos$PROCEDIMIENTO == "ANGIOPLASTIA"),]),nrow(tresVasos[which(tresVasos$PROCEDIMIENTO == "CIRUGIA"),]),nrow(tresVasos[which(tresVasos$PROCEDIMIENTO == "ANGIOPLASTIA"),]))
dataVasosProced <- data.frame(procedimiento,coronopatia,values)
grafico1<-ggplot(dataVasosProced, aes(fill=procedimiento, y=values ,x=procedimiento)) + 
  geom_bar(position="dodge", stat="identity") +
  scale_fill_viridis(discrete = T, option = "E") +
  facet_wrap(~coronopatia,nrow = 1) +
  theme_ipsum() +
  theme(legend.position="none") +
  scale_fill_manual(values = c("#6495ED","#5f9ea0"))+
  xlab("")
```
```{r fig.width=8, fig.height=6, fig.align='center',echo=FALSE,warning=FALSE,cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
grafico1
```
<font size="2">This graph visually represents how most of the patients with 1 and 2 affected vessels end up in angioplasty, while 3 vessels patients are almost equally distributed between angioplasty and surgery.</font>
<p>&nbsp;</p>



<font size="5.5"> **Procedure Frequency by Risk Factor** </font>

```{r, include= FALSE,results='asis'}
riskCirugia <- dataset[which(dataset$PROCEDIMIENTO == "CIRUGIA"),] 
frecRiesgoCirugia<- c(as.numeric(length(factor(which(riskCirugia$EPOC == 1)))), as.numeric(length(factor(which(riskCirugia$DIABETES == 1)))), as.numeric(length(factor(which(riskCirugia$HIPERTENSION.PULMONAR == 1)))), as.numeric(length(factor(which(riskCirugia$OBESIDAD.MORBIDA == 1)))),  as.numeric(length(factor(which(riskCirugia$EDAD >65)))), as.numeric(length(factor(which(riskCirugia$FEY <50))))) 
cols<- c("EPOC", "Diabetes", "Lung Hypertension", "Morbid Obesity", "Age above 65", "FEY less than 50")
rows<- c("Cirugia", "Angioplastia")
riskATC <- dataset[which(dataset$PROCEDIMIENTO == "ANGIOPLASTIA"),]
frecRiesgoATC<- c(as.numeric(length(factor(which(riskATC$EPOC == 1)))), as.numeric(length(factor(which(riskATC$DIABETES == 1)))), as.numeric(length(factor(which(riskATC$HIPERTENSION.PULMONAR == 1)))), as.numeric(length(factor(which(riskATC$OBESIDAD.MORBIDA == 1)))),  as.numeric(length(factor(which(riskATC$EDAD >65)))), as.numeric(length(factor(which(riskATC$FEY <50))))) 

procedRisksData<-data.frame(frecRiesgoCirugia,frecRiesgoATC, row.names = cols)
colnames(procedRisksData)[1]<-"Surgery"
colnames(procedRisksData)[2]<-"Angioplasty"

procedRisksData
```
```{r fig.width=10, fig.height=5, fig.align='center',echo=FALSE,warning=FALSE}
barplot(t(as.matrix(procedRisksData)),beside=TRUE,xlab="Risk Factors", ylab="Frequency", col=c("#cd5c5c", "#6495ed"),
        legend.text = colnames(procedRisksData),args.legend=list(cex=0.75,x = "topright"))
```
<font size="2">The aim of this graph is to show the frequency of procedures according to the main risk factors in the dataset. It can easily be seen how diabetic patients  and the ones above 65 years old end up in some kind of procedure, mainly angioplasty.</font">

# Predictor 

<font size="5.5"> **Team Members** </font>

#### [Santiago Encina](mailto:santiago.encina@ing.austral.edu.ar)
#### [Matias Coulougian](mailto:matias.coulougian@ing.austral.edu.ar)
#### [Ricardo Lopez Bonaguro](mailto:ricardo.lopez@ing.austral.edu.ar)
#### [Camila Catalini](mailto:camila.catalini@ing.austral.edu.ar)
#### [Alejo Demitropulos](mailto:alejo.demitropulos@ing.austral.edu.ar)

## {.unnumbered}