---
title: "Our Second R-Markdown"
author: "Grupo 8"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---
# First analysis and graphs test from the dataset

## Sex Distribution 
```{r, include = FALSE, results='asis'}
library(plotrix)
library(plotly)
library(ggplot2)
library(ggridges)
library(dplyr)
library(hrbrthemes)

dataset <- read.csv("C:/Users/54112/Desktop/TpR/dataset.csv", sep=";")
sexo <- factor(dataset$SEXO)
a <- table(sexo)
porcentajeHyM <- prop.table(a)
porcentajeHyM <- porcentajeHyM*100
porcentajeHyM <- round(porcentajeHyM[c(1,2)], 2)
percentageVector <- as.numeric(porcentajeHyM)
women <- percentageVector[1]
men <- percentageVector[2]
diab <- dataset[which(dataset$DIABETES == "1"),]
diabPerc <- round((nrow(diab)/nrow(dataset))*100)
notDiabPerc <- 100 - diabPerc

```
#### Code: 
```{r fig.width=6, fig.height=6, fig.align='center'}
pie3D(c(women,men),main = "Sex Distribution", 
      labels = c(paste(women,"%"),paste(men,"%")),
      col = c("turquoise","lightblue"),labelcex = 1, 
      explode = 0.1,start = 3 ,theta = 1.2, mar = c(6,6,6,6))
legend(-0.9,-0.8,c("Women","Men"),fill = c("turquoise","lightblue"))
```
In this graph we can observe the sex distribution. There is a much higher percentage of men in the dataset.

## Diabetic Pacients Distribution

#### Code: 
```{r fig.width=6, fig.height=6, fig.align='center'}
pie3D(c(diabPerc,notDiabPerc),main = "Diabetics Distribution", 
      labels = c(paste(diabPerc,"%"),paste(notDiabPerc,"%")),
      col = c("red","lightgreen"),labelcex = 1, explode = 0.1,start = 2.9 ,
      theta = 1.1, mar = c(6,6,6,6))
legend(-1.1,1.15,c("Diabetic","Non Diabetic"),fill = c("red","lightgreen"))
```

## A New Output Format

`html_pretty` in the [**prettydoc**](https://github.com/yixuan/prettydoc/)
package is a new output format for creating HTML documents from R Markdown files.
`html_pretty` is more lightweight compared to `html_document`, and is more
stylish than `html_vignette` when creating package vignettes.

## New Styles

`html_pretty` currently supports three page themes (`cayman`, `tactile` and
`architect`), and two syntax highlight styles (`github` and `vignette`).
The theme and highlight styles can be specified in the document metadata,
for example:

```yaml
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
```

## Your Familiar Tools

Feel free to use the `knitr` infrastructure with dozens of tunable options in
your document.

```{r fig.width=6, fig.height=6, fig.align='center'}
set.seed(123)
n <- 1000
x1  <- matrix(rnorm(n), ncol = 2)
x2  <- matrix(rnorm(n, mean = 3, sd = 1.5), ncol = 2)
x   <- rbind(x1, x2)
head(x)
smoothScatter(x, xlab = "x1", ylab = "x2")
```

You can include code snippets of languages other than R, but note that
the block header has no curly brackets around the language name.

```cpp
// [[Rcpp::export]]
NumericVector timesTwo(NumericVector x) {
    return x * 2;
}
```

You can also write math expressions, e.g. $Y = X\beta + \epsilon$,
footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(iris, 10))
```

## Stay Tuned

Please visit the [development page](https://github.com/yixuan/prettydoc/) of the 
`prettydoc` package for latest updates and news. Comments, bug reports and
pull requests are always welcome.
